---
title: "plotly"
author: "Hendrilalaina"
date: "2025-03-31"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(dplyr))
```

# Introduction

We will use data from World Happiness Report. The World Happiness Report is a landmark survey of the state of global happiness.

# Data Preparation

The data can be downloaded from (<https:www.kaggle.com/unsdsn/world-happiness/>)

```{r}
happiness_2015 <- read.csv("../RUltimateMaterial-main/data/2015.csv") %>%
  mutate(year = 2015)
happiness_2016 <- read.csv("../RUltimateMaterial-main/data/2016.csv") %>%
  mutate(year = 2016)
happiness_2017 <- read.csv("../RUltimateMaterial-main/data/2017.csv") %>%
  mutate(year = 2017)

happiness <- happiness_2015 %>%
  bind_rows(happiness_2016) %>%
  bind_rows(happiness_2017) %>%
  select(Country, `Happiness.Score`, `Economy..GDP.per.Capita.`, `Health..Life.Expectancy.`, Freedom, year)
happiness %>% glimpse()
```
Rename columns

```{r}
happiness_mod <- happiness %>%
  rename(happiness_score = Happiness.Score) %>%
  rename(economy_gdp = Economy..GDP.per.Capita.) %>%
  rename(health_life_exp = Health..Life.Expectancy.)
```

# Visualization

## Complet Plot

We create a complex visualization (5 dimensional), which has these properties: 
- x axis refers to GDP (an economical value) 
- y axis refers to life expectancy (a health value) 
- size refers to freedom - color refers to happiness score 
- text refers to country - slider refers to the year

```{r}
p1 <- happiness_mod %>%
  filter(year == 2015) %>%
  plot_ly(x = ~economy_gdp,
          y = ~health_life_exp,
          size = ~Freedom^5,
          color = ~happiness_score,
          text = ~Country,
          hoverinfo = "text") %>%
  add_markers() %>%
  colorbar(title = "Happiness Score") %>%
  layout(title = "Happiness 2015",
         xaxis = list(title = "Economy (repr. by GDP)"),
         yaxis = list(title = "Healthy repr. by Life Expectancy"))
```
## Subplots

```{r}
p2 <- happiness_mod %>%
  filter(year == 2016) %>%
  plot_ly(x = ~economy_gdp,
          y = ~health_life_exp,
          size = ~Freedom^5,
          color = ~happiness_score,
          text = ~Country,
          hoverinfo = "text") %>%
  add_markers() %>%
  colorbar(title = "Happiness Score") %>%
  layout(title = "Happiness 2015/2016",
         xaxis = list(title = "Economy (repr. by GDP)"),
         yaxis = list(title = "Healthy repr. by Life Expectancy"))
subplot(titleX = TRUE, shareY = TRUE, p1, p2)
```

