---
title: "dygraphs"
author: "Hendrilalaina"
date: "2025-04-07"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_folding: show
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will directly download gold and silver price data to R with _quantmod_ package and create a dynamic plot with _dygraphs_.
* Objectives: basic use of _quantmod_ and _dygraphs_.
* Requirements: R Basics

# Data Download and Understanding

We begin loading required packages: _quantmod_ and _dygraphs_. You can install _install.packages()_ from CRAN. As of date of writing you need to install _dygraphs_ directly from github. For this you can use the function call below. After installing it once, you can load it.

```{r}
suppressPackageStartupMessages(library(quantmod))
suppressPackageStartupMessages(library(dygraphs))
```

Now we download data. This is done with _getMetals()_ from _quantmod_ package. Metals are passed as a vector with ticker names for metals. "From" is supposed to be the start of the period, but only the last 5 years are provided as we will see. So this parameter is partially ignored, but if you don't set it a much shorter period is provided.

For each of these metals an xts-object is stored. These objects are combined with cbind to one single sct-object.

```{r download_data, cache=T}
quantmod::getMetals(Metals = c('XAU', 'XAG', 'PALL', 'PL'),
                    from = "2001-01-01",
                    to = Sys.Date())
metals <- cbind(XAGUSD, XAUUSD, XPDUSD, XPTUSD)
```

# Dynamic Plot

Let's skip the first line. We will come to that later.

Plot is created with _dygraphs()_ function. Similar to _ggplot()_ you start with basic information on dataframe, and the put more aesthetics on top. Most importantly "data" is passed. It has to be a timeseries xts-object. Some general information title, x- and y- label can be passed as well. More definitions on how the plot should look and feel can be added with piping "%>%".

With _dyRebase()_ all prices are referenced to 100 at beginning of period.
_dyRangeSelector()_ defines highlighted area. Plot can be extended to other periods with slider at the bottom of the page.

```{r}
dy_window <- c("2024-06-01", "2025-04-01")
dygraph(data = metals,
        xlab = "Time",
        ylab = "Gold, Silver, Platinum, and Palladium Performance") %>%
  dyRebase(value = 100) %>%
  dyRangeSelector(dateWindow = dy_window)

```

This plot is very dynamic. You can modify time range by changing start of end, or moving with a fixed period. Specific areas can be zoomed in by clicking on the graph. If you want to unzoom just double-click in the graph. Prices are always shown for current position at the top of the plot.

All metals increased prices since September 2018, Palladium by more than 60%. But this picture changes when period is extended to the last 5 years.

# Linked Plots

```{r}
dygraph(XAUUSD, group = "linked") %>%
  dyRangeSelector(dateWindow = dy_window)

dygraph(XAUUSD, group = "linked") %>%
  dyRangeSelector(dateWindow = dy_window)
```

# More Informations

There is more to discover. Take a look at these websites to learn more about financial modeling with _quantmod_ or dynamic timseries plotting with _dygraphs_.

* Quantmod Package <http://www.quantmod.com>
* Dygraphs Package <https://rstudio.github.io/dygraphs/>
